---
title: "2023 PSRC Report Tables"
author: "James Clark"
date: "`r Sys.Date()`"
output:
  word_document
---


```{r, setup, echo = FALSE, warning = FALSE}

# Setup=========================================================================

knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)

# remotes::install_github('RSGInc/travelSurveyTools')
library(travelSurveyTools)
library(data.table)
library(stringr)
library(janitor)
library(readxl)
library(flextable)

user = Sys.info()['user']

data_dir = str_glue(
  "C:/Users/{user}/Resource Systems Group, Inc/Transportation MR - Documents/",
  "PSRC Survey Program/210252_PSRC_HTS/",
  "2023 Puget Sound Travel Study/5.Deliverables/7_Data&Documentation/",
  "0_Data_Inputs_From_PSRC/from_psrc_20240201"
)

hh = fread(file.path(data_dir, 'Household.csv'))
person = fread(file.path(data_dir, 'Person.csv'))
day = fread(file.path(data_dir, 'Day.csv'))
trip = fread(file.path(data_dir, 'Trip.csv'))

# load codebook
codebook_path = str_glue(
  "C:/Users/{user}/Resource Systems Group, Inc/Transportation MR - Documents/",
  "PSRC Survey Program/210252_PSRC_HTS/Internal/3.DataAnalysis/1.Data/Codebook/",
  "PSRC_Combined_Codebook_2023_08162023_RSG.xlsx"
)

variable_list = read_excel(codebook_path, sheet = 'variable_list_2023')
value_labels = read_excel(codebook_path, sheet = 'value_labels_2023')
value_crosswalk = read_excel(codebook_path, sheet = 'value_crosswalk_2023')

setDT(variable_list)
setDT(value_labels)
setDT(value_crosswalk)

wt_dir = str_glue(
  "C:/Users/{user}/Resource Systems Group, Inc/Transportation MR - Documents/",
  "PSRC Survey Program/210252_PSRC_HTS/2023 Puget Sound Travel Study/",
  "5.Deliverables/9_Weighting/2_RSG_Outputs"
)

# Read in weights
hh_weights = fread(file.path(wt_dir, 'hh_weights.csv'))
person_weights = fread(file.path(wt_dir, 'person_weights.csv'))
day_weights = fread(file.path(wt_dir, 'day_weights.csv'))
trip_weights = fread(file.path(wt_dir, 'trip_weights.csv'))


```


```{r data_prep, echo = FALSE, warning = FALSE}

# Prepare Data for tables=======================================================

# Prepare codebook--------------------------------------------------------------

# rename columns to match table names

old_table_names = names(variable_list)[str_detect(names(variable_list), '_final')]
new_table_names = str_remove(old_table_names, '_final')

setnames(variable_list, old_table_names, new_table_names)

# Rename other columns to match expected input for travelSurveyTools

old_names = c('psrc_variable', 'description_2023', 'data_type_2023', 'var_logic')
new_names = c('variable', 'description', 'data_type', 'logic')

setnames(variable_list, old_names, new_names)

# Only keep the important variables
keep_vars = c(new_table_names, new_names)

variable_list = variable_list[, ..keep_vars]

# Find checkbox variables i.e. have "-- " in the description

variable_list[, is_checkbox := 1 * str_detect(description, "--")]

# Create shared_name variable
variable_list[, shared_name :=
                ifelse(is_checkbox == 1,
                       sub("_[^_]*$", "", variable),
                       variable
                )]

# assigned tables, weights, and ids to lists to feed into functions

# Append the weights------------------------------------------------------------

# drop old weights (all placeholder 1s)
hh[, hh_weight_2023 := NULL]
person[, person_weight := NULL]
day[, day_weight_2023 := NULL]
trip[, trip_weight := NULL]

# make all ids characters
hh[, hhid := as.character(hhid)]
hh_weights[, hh_id := as.character(hh_id)]

person[, person_id := as.character(person_id)]
person[, hhid := as.character(hhid)]
person_weights[, person_id := as.character(person_id)]

day[, day_id := as.character(day_id)]
day[, hhid := as.character(hhid)]
day[, person_id := as.character(person_id)]
day_weights[, day_id := as.character(day_id)]

trip[, tripid := as.character(tripid)]
trip[, hhid := as.character(hhid)]
trip[, person_id := as.character(person_id)]
trip[, day_id := as.character(day_id)]
trip_weights[, trip_id := as.character(trip_id)]

# attach new weights
hh[
  hh_weights,
  hh_weight_2023 := i.hh_weight,
  on = c('hhid' = 'hh_id')
]

person[
  person_weights,
  person_weight := i.person_weight,
  on = 'person_id'
]

day[
  day_weights,
  day_weight_2023 := i.day_weight,
  on = 'day_id'
]

trip[
  trip_weights,
  trip_weight := i.trip_weight,
  on = c('tripid' = 'trip_id')
]

#check the weights
stopifnot(sum(hh$hh_weight_2023, na.rm = T) == sum(hh_weights$hh_weight))
stopifnot(sum(person$person_weight, na.rm = T) == sum(person_weights$person_weight))
stopifnot(sum(day$day_weight_2023, na.rm = T) == sum(day_weights$day_weight))
stopifnot(sum(trip$trip_weight, na.rm = T) == sum(trip_weights$trip_weight))

tables = list('hh' = hh,
              'person' = person,
              'day' = day,
              'trip' = trip)

# drop households that are no longer complete after re-calculating completes
tables = hts_filter_data(tables, 
                         hh[!is.na(hh_weight_2023), hhid],
                         'hhid')

ids = c('hhid', 'person_id', 'day_id', 'tripid')
wts = c('hh_weight_2023', 'person_weight', 'day_weight_2023', 'trip_weight')

```

```{r define functions, echo = FALSE, warning = FALSE}

# Create functions to make tables using travelSurveyTools-----------------------

# functions makes flextable with weighted/unweighted counts + percents
make_reporting_table = function(summarize_var,
                                variable_list,
                                tables,
                                ids,
                                wts,
                                wtname,
                                label,
                                value_labels = NULL,
                                shared = FALSE){
  
  # create input for hts_summary
  prep = hts_prep_variable(
    summarize_var = summarize_var,
    variables_dt = variable_list,
    data = tables,
    id_cols = ids,
    wt_cols = wts
  )
  
  #edit parameters for checkbox variables
  if (shared){
    
    checkbox_yes = 1
    summarize_type = 'checkbox'
    
  } else {
    
    checkbox_yes = NULL
    summarize_type = 'categorical'
    
  }
  
  # create summary
  output = hts_summary(prep$cat,
                       summarize_var = summarize_var,
                       summarize_vartype = summarize_type,
                       id_cols = ids,
                       wtname = wtname,
                       checkbox_yesval = checkbox_yes)
  
  
  unwtd = output$summary$unwtd
  wtd = output$summary$wtd
  
  # format outputs for table
  setnames(unwtd, 'count', 'Unweighted Count')
  
  unwtd[, `Unweighted %` := round(prop * 100, 2)]
  
  setnames(wtd, 'est', 'Weighted Count')
  
  wtd[, `Weighted Count` := round(`Weighted Count`)]
  
  wtd[, `Weighted %` := round(prop * 100, 2)]
  
  tab = merge(unwtd, wtd, all.x = TRUE, by = summarize_var)
  
  #subset variables
  keep_vars = c(summarize_var, 'Unweighted Count', 'Weighted Count', 'Unweighted %', 'Weighted %')
  
  tab = tab[, ..keep_vars]
  
  # label data
  if (label){
    
    tab = factorize_df(
      tab,
      value_labels,
      variable_colname = 'psrc_variable',
      value_colname = 'psrc_value',
      value_order_colname = 'psrc_value',
      value_label_colname = "final_label")
    
  }
  
  # return as a formated flextable
  tab %>% 
    adorn_totals(where = 'row') %>%
    flextable() %>% 
    colformat_double(j = 4:5, suffix = '%')
  
}

# function to make table for trip rates-----------------------------------------
make_triprate_table = function(summarize_by,
                               variable_list,
                               tables,
                               ids,
                               wts,
                               label = FALSE
){
  
  # Prep unweighted data
  triprate_prep = hts_prep_triprate(summarize_by = summarize_by,
                                    variables_dt = variable_list,
                                    hts_data = tables,
                                    ids = ids,
                                    wts = wts,
                                    remove_outliers = FALSE,
                                    weighted = FALSE)
  
  # get unwtd counts
  unwtd_counts = triprate_prep$num[, .(`Unweighted Days` =  .N,
                                       `Unweighted Trips` = sum(num_trips_unwtd)),
                                   summarize_by]
  
  
  # get unwtd trip rates
  output = hts_summary(triprate_prep$num,
                       summarize_var = 'num_trips_unwtd',
                       summarize_by = summarize_by,
                       summarize_vartype = 'numeric',
                       id_cols = ids,
                       wtname = 'day_weight_2023')
  
  unwtd_triprate = output$summary$unwtd[, 
                                        .(
                                          summarize_by = get(summarize_by),
                                          'Unweighted Trip Rate' =  round(mean, 2)
                                        )
  ]
  
  
  unwtd_tab = merge(unwtd_counts, unwtd_triprate, by.x = summarize_by, by.y = 'summarize_by')
  
  # prep weighted data
  triprate_prep = hts_prep_triprate(summarize_by = summarize_by,
                                    variables_dt = variable_list,
                                    hts_data = tables,
                                    ids = ids,
                                    wts = wts,
                                    remove_outliers = FALSE,
                                    weighted = TRUE)
  
  # get weighted counts
  wtd_counts = triprate_prep$num[, .(`Weighted Days` = round(sum(day_weight_2023)),
                                     `Weighted Trips` = round(sum(num_trips_wtd))),
                                 summarize_by]
  
  # get weighted trip rates
  output = hts_summary(triprate_prep$num,
                       summarize_var = 'num_trips_wtd',
                       summarize_by = summarize_by,
                       summarize_vartype = 'numeric',
                       id_cols = ids,
                       wtname = 'day_weight_2023')
  
  wtd_triprate = output$summary$wtd[, 
                                    .(
                                      summarize_by = get(summarize_by),
                                      'Weighted Trip Rate' = round(mean, 2)
                                    )
  ]
  
  
  wtd_tab = merge(wtd_counts, wtd_triprate, by.x = summarize_by, by.y = 'summarize_by')
  
  #combined weighted and unweighted
  tab = merge(unwtd_tab, wtd_tab, by = summarize_by)
  
  tab = tab[!get(summarize_by) %in%  c('NULL', 'NA')]
  
  # label and order
  if (label){
    
    tab = factorize_df(
      tab,
      value_labels,
      variable_colname = 'psrc_variable',
      value_colname = 'psrc_value',
      value_order_colname = 'psrc_value',
      value_label_colname = "final_label")
    
  }
  
  # Add total row
  total = data.table(
    summarize_by = 'Total',
    'Unweighted Days' = sum(tab$`Unweighted Days`),
    'Unweighted Trips' = sum(tab$`Unweighted Trips`),
    'Weighted Days' = sum(tab$`Weighted Days`),
    'Weighted Trips' = sum(tab$`Weighted Trips`)
  )
  
  setnames(total, 'summarize_by', summarize_by)
  
  total[, `Unweighted Trip Rate` := `Unweighted Trips` / `Unweighted Days`]
  total[, `Weighted Trip Rate` := `Weighted Trips` / `Weighted Days`]
  
  tab = rbind(tab, total)
  
  # round trip rates
  tab[, `Unweighted Trip Rate` := round(`Unweighted Trip Rate`, 2)]
  tab[, `Weighted Trip Rate` := round(`Weighted Trip Rate`, 2)]
  
  setcolorder(tab, 
              c(summarize_by,
                'Unweighted Days',
                'Unweighted Trips',
                'Weighted Days',
                'Weighted Trips',
                'Unweighted Trip Rate',
                'Weighted Trip Rate')
  )
  
  #return as flextable
  return(tab %>% flextable())
  
}


```


# Demographics by participation group-------------------------------------------

# Demographics

# Counts by diary platform

```{r}

rmove_ids = tables$hh[diary_platform == 'rmove', hhid]
browser_ids = tables$hh[diary_platform == 'browser', hhid]
call_ids = tables$hh[diary_platform == 'call', hhid]


diary_platform_counts = data.table(
  table = c('hh', 'person', 'trip'),
  rmove_counts = c(
    tables$hh[hhid %in% rmove_ids, .N],
    tables$person[hhid %in% rmove_ids, .N],
    tables$trip[hhid %in% rmove_ids, .N]
  ),
  browser_counts = c(
    tables$hh[hhid %in% browser_ids, .N],
    tables$person[hhid %in% browser_ids, .N],
    tables$trip[hhid %in% browser_ids, .N]
  ),
  call_counts = c(
    tables$hh[hhid %in% call_ids, .N],
    tables$person[hhid %in% call_ids, .N],
    tables$trip[hhid %in% call_ids, .N]
  )
)


diary_platform_counts %>% 
  adorn_totals(where = 'col') %>% 
  flextable()

```


# TABLE 11: PARTICIPANT AGE BY PARTICIPATION MODE

```{r table 11, echo=FALSE}

prep = hts_prep_variable(summarize_var = 'age',
                         summarize_by = 'diary_platform',
                         variables_dt = variable_list,
                         data = tables,
                         id_cols = ids,
                         wt_cols = wts)


output = hts_summary(prep$cat,
                     summarize_var = 'age',
                     summarize_by = 'diary_platform',
                     summarize_vartype = 'categorical',
                     id_cols = ids,
                     wtname = 'person_weight')

tab = output$summary$wtd

tab$pct = round(tab$prop * 100, 2)

tab = factorize_df(
  tab,
  value_labels,
  variable_colname = 'psrc_variable',
  value_colname = 'psrc_value',
  value_order_colname = 'psrc_value',
  value_label_colname = "final_label")

dcast(tab, age ~ diary_platform, value.var = 'pct') %>% 
  adorn_totals() %>% 
  flextable() %>%
  colformat_double(suffix = '%')


```

# TABLE 12: PARTICIPANT HOUSEHOLD INCOME BY PARTICIPATION MODE

```{r table 12, echo=FALSE}
prep = hts_prep_variable(summarize_var = 'hhincome_broad',
                         summarize_by = 'diary_platform',
                         variables_dt = variable_list,
                         data = tables,
                         id_cols = ids,
                         wt_cols = wts)


output = hts_summary(prep$cat,
                     summarize_var = 'hhincome_broad',
                     summarize_by = 'diary_platform',
                     summarize_vartype = 'categorical',
                     id_cols = ids,
                     wtname = 'hh_weight_2023')

tab = output$summary$wtd

tab$pct = round(tab$prop * 100, 2)

tab = factorize_df(
  tab,
  value_labels,
  variable_colname = 'psrc_variable',
  value_colname = 'psrc_value',
  value_order_colname = 'psrc_value',
  value_label_colname = "final_label")

dcast(tab, hhincome_broad ~ diary_platform, value.var = 'pct') %>% 
  adorn_totals() %>% 
  flextable() %>%
  colformat_double(suffix = '%')

```

# TABLE 13: PARTICIPANT HOUSEHOLD SIZE BY PARTICIPATION MODE

```{r table 13, echo=FALSE}

# create a binned variable for household size
tables$hh[, hhsize_binned := ifelse(hhsize > 5, 5, hhsize)]

# Write new variable to variable list
variable_list = rbind(
  variable_list,
  data.table(
    variable = 'hhsize_binned',
    hh = 1,
    person = 0,
    vehicle = 0,
    day = 0,
    trip = 0,
    location = 0,
    description = 'binned household size',
    data_type = 'integer/categorical',
    logic = NA,
    is_checkbox = 0,
    shared_name = 'hhsize_binned'
  )
)





prep = hts_prep_variable(summarize_var = 'hhsize_binned',
                         summarize_by = 'diary_platform',
                         variables_dt = variable_list,
                         data = tables,
                         id_cols = ids,
                         wt_cols = wts)


output = hts_summary(prep$cat,
                     summarize_var = 'hhsize_binned',
                     summarize_by = 'diary_platform',
                     summarize_vartype = 'categorical',
                     id_cols = ids,
                     wtname = 'hh_weight_2023')

tab = output$summary$wtd

tab$pct = round(tab$prop * 100, 2)

tab = factorize_df(
  tab,
  value_labels,
  variable_colname = 'psrc_variable',
  value_colname = 'psrc_value',
  value_order_colname = 'psrc_value',
  value_label_colname = "final_label")

dcast(tab, hhsize_binned ~ diary_platform, value.var = 'pct') %>% 
  adorn_totals() %>% 
  flextable() %>%
  colformat_double(suffix = '%')

```

# TABLE 14: PARTICIPANT RACE BY PARTICIPATION MODE

```{r table 14, echo=FALSE}

# Recode person table so each person is in only one race category

# Exclude Hispanic because it was asked in ethnicity
race_vars = variable_list[shared_name == 'race' & variable != 'race_hisp', variable]

#count responses for race
tables$person[, num_races := rowSums(tables$person[, ..race_vars] == 1)]

ethnicity_vars = variable_list[shared_name == 'ethnicity', variable]

# count responses for ethnicity
tables$person[, num_ethnicities := rowSums(tables$person[, ..ethnicity_vars] == 1)]

# code race_hisp
tables$person[, race_hisp := as.integer(ifelse(race_white == 1 & ethnicity_1 == 0, 1, 0))]

tables$person[ race_hisp == 1, race_white := 0]

# Create variables for multiple races
tables$person[, race_multiple := as.integer(ifelse(num_races > 1 | num_ethnicities > 1, 1, 0))]



tables$person[race_multiple == 1, `:=` 
              (race_afam = 0,
                race_aiak = 0,
                race_asian = 0,
                race_hapi = 0,
                race_noanswer = 0,
                race_other = 0,
                race_white = 0,
                race_hisp = 0)]

variable_list = rbind(
  variable_list,
  data.table(
    variable = 'race_multiple',
    hh = 1,
    person = 0,
    vehicle = 0,
    day = 0,
    trip = 0,
    location = 0,
    description = 'Race -- Multiple races or ethnicities',
    data_type = 'integer/categorical',
    logic = NA,
    is_checkbox = 0,
    shared_name = 'race'
  )
)


prep = hts_prep_variable(summarize_var = 'race',
                         summarize_by = 'diary_platform',
                         variables_dt = variable_list,
                         data = tables,
                         id_cols = ids,
                         wt_cols = wts)


output = hts_summary(prep$cat,
                     summarize_var = 'race',
                     summarize_by = 'diary_platform',
                     summarize_vartype = 'checkbox',
                     id_cols = ids,
                     wtname = 'person_weight',
                     checkbox_yesval = 1)

tab = output$summary$wtd

tab$pct = round(tab$prop * 100, 2)

tab = factorize_df(
  tab,
  value_labels,
  variable_colname = 'psrc_variable',
  value_colname = 'psrc_value',
  value_order_colname = 'psrc_value',
  value_label_colname = "final_label")

dcast(tab, race ~ diary_platform, value.var = 'pct') %>% 
  adorn_totals() %>% 
  flextable() %>%
  colformat_double(suffix = '%')


```


# TABLE 15: TABLE 15: HOUSEHOLD SIZE (WEIGHTED)

```{r table 15, echo = FALSE, warning = FALSE}

make_reporting_table(summarize_var = 'hhsize_binned',
                     variable_list = variable_list,
                     tables = tables,
                     ids = ids,
                     wts = wts,
                     wtname = 'hh_weight_2023',
                     label = FALSE)


```

# TABLE 16: HOUSEHOLD INCOME (WEIGHTED)

```{r table 16, echo = FALSE, warning = FALSE}

make_reporting_table(summarize_var = 'hhincome_broad',
                     variable_list = variable_list,
                     tables = tables,
                     ids = ids,
                     wts = wts,
                     wtname = 'hh_weight_2023',
                     label = TRUE,
                     value_labels = value_labels)

```

# TABLE 17: HOUSEHOLD VEHICLES (WEIGHTED)

```{r table 17, echo = FALSE, warning = FALSE}

# create a binned variable for num_vehicles
tables$hh[, vehicle_count_binned := ifelse(vehicle_count > 3, 3, vehicle_count)]

# Write new variable to variable list
variable_list = rbind(
  variable_list,
  data.table(
    variable = 'vehicle_count_binned',
    hh = 1,
    person = 0,
    vehicle = 0,
    day = 0,
    trip = 0,
    location = 0,
    description = 'binned number of vehicles',
    data_type = 'integer/categorical',
    logic = NA,
    is_checkbox = 0,
    shared_name = 'vehicle_count_binned'
  )
)



make_reporting_table(summarize_var = 'vehicle_count_binned',
                     variable_list = variable_list,
                     tables = tables,
                     ids = ids,
                     wts = wts,
                     wtname = 'hh_weight_2023',
                     label = FALSE)

```

# TABLE 18: PERSON AGE (WEIGHTED)

```{r table 18, echo = FALSE, warning = FALSE}
make_reporting_table(summarize_var = 'age',
                     variable_list = variable_list,
                     tables = tables,
                     ids = ids,
                     wts = wts,
                     wtname = 'person_weight',
                     label = TRUE,
                     value_labels = value_labels)

```

# TABLE 19: PERSON GENDER (WEIGHTED)

```{r table 19, echo = FALSE, warning = FALSE}
make_reporting_table(summarize_var = 'gender',
                     variable_list = variable_list,
                     tables = tables,
                     ids = ids,
                     wts = wts,
                     wtname = 'person_weight',
                     label = TRUE,
                     value_labels = value_labels)

```

# TABLE 20: PERSON RACE OR ETHNICITY

```{r table 20, echo = FALSE, warning = FALSE}

make_reporting_table(summarize_var = 'race',
                     variable_list = variable_list,
                     tables = tables,
                     ids = ids,
                     wts = wts,
                     wtname = 'person_weight',
                     label = TRUE,
                     value_labels = value_labels,
                     shared = TRUE)

```

# TABLE 21: PERSON STUDENT STATUS 

```{r table 21, echo = FALSE, warning = FALSE}

# get values from codebook to classify
part_time = value_labels[
  psrc_variable == 'student' &
    str_detect(final_label, 'Part-time'),
  psrc_value]

full_time = value_labels[
  psrc_variable == 'student' &
    str_detect(final_label, 'Full-time'),
  psrc_value]

# make a binned student variable
tables$person[, student_binned := fcase(
  student %in% full_time, 'Full-time',
  student %in% part_time, 'Part-time',
  default = 'Not a student'
)]

# add to variable list
variable_list = rbind(
  variable_list,
  data.table(
    variable = 'student_binned',
    hh = 0,
    person = 1,
    vehicle = 0,
    day = 0,
    trip = 0,
    location = 0,
    description = 'Student values binned',
    data_type = 'categorical',
    logic = NA,
    is_checkbox = 0,
    shared_name = 'student_binned'
  )
)


make_reporting_table(summarize_var = 'student_binned',
                     variable_list = variable_list,
                     tables = tables,
                     ids = ids,
                     wts = wts,
                     wtname = 'person_weight',
                     label = TRUE,
                     value_labels = value_labels)

```

# TABLE 22: PERSON EMPLOYMENT STATUS (AGE 18+)

```{r table 22, echo = FALSE, warning = FALSE}

make_reporting_table(summarize_var = 'employment',
                     variable_list = variable_list,
                     tables = tables,
                     ids = ids,
                     wts = wts,
                     wtname = 'person_weight',
                     label = TRUE,
                     value_labels = value_labels)

```

# TABLE 23: PERSON SEXUALITY

```{r table 23, echo = FALSE, warning = FALSE}

make_reporting_table(summarize_var = 'sexuality',
                     variable_list = variable_list,
                     tables = tables,
                     ids = ids,
                     wts = wts,
                     wtname = 'person_weight',
                     label = TRUE,
                     value_labels = value_labels)

```

# Trip Rates--------------------------------------------------------------------

# TABLE 24: PERSON TRIP RATE, BY DAY OF WEEK 

```{r table 24, echo = FALSE, warning = FALSE}

make_triprate_table(summarize_by = 'travel_dow',
                    variable_list = variable_list,
                    tables = tables,
                    ids = ids,
                    wts = wts,
                    label = TRUE)


```


# TABLE 25: PERSON TRIP RATE, BY HOUSEHOLD INCOME

```{r table 25, echo = FALSE, warning = FALSE}

make_triprate_table(summarize_by = 'hhincome_broad',
                    variable_list = variable_list,
                    tables = tables,
                    ids = ids,
                    wts = wts,
                    label = TRUE)

```

# TABLE 26: PERSON TRIP RATE, BY AGE GROUP 

```{r table 26, echo = FALSE, warning = FALSE}

make_triprate_table(summarize_by = 'age',
                    variable_list = variable_list,
                    tables = tables,
                    ids = ids,
                    wts = wts,
                    label = TRUE)

```

# TABLE 27: PERSON TRIP RATE, BY EMPLOYMENT STATUS (AGE 18+)

```{r table 27, echo = FALSE, warning = FALSE}

make_triprate_table(summarize_by = 'employment',
                    variable_list = variable_list,
                    tables = tables,
                    ids = ids,
                    wts = wts,
                    label = TRUE)

```

# TABLE 28: PERSON TRIP RATE, BY TRAVEL MODE 

```{r table 28, echo = FALSE, warning = FALSE}

make_triprate_table(summarize_by = 'mode_type',
                    variable_list = variable_list,
                    tables = tables,
                    ids = ids,
                    wts = wts,
                    label = TRUE)


```

# TABLE 29: PERSON TRIP RATE, BY TRIP PURPOSE 

```{r table 29, echo = FALSE, warning = FALSE}

make_triprate_table(summarize_by = 'dest_purpose_cat',
                    variable_list = variable_list,
                    tables = tables,
                    ids = ids,
                    wts = wts,
                    label = TRUE)


```

# Percent of trips by travel mode-----------------------------------------------

# TABLE 30: TRAVEL MODE, BY HOUSEHOLD INCOME (WEIGHTED)

```{r table 30, echo = FALSE, warning = FALSE}

prep = hts_prep_variable(summarize_var = 'mode_type',
                         summarize_by = 'hhincome_broad',
                         variables_dt = variable_list,
                         data = tables,
                         id_cols = ids,
                         wt_cols = wts,
                         missing_values = c("995", "NULL"))


output = hts_summary(prep$cat,
                     summarize_var = 'mode_type',
                     summarize_by = 'hhincome_broad',
                     summarize_vartype = 'categorical',
                     id_cols = ids,
                     wtname = 'trip_weight')

tab = output$summary$wtd

tab$pct = round(tab$prop * 100, 2)

# tab = tab[mode_type != 'NULL']

tab = factorize_df(
  tab,
  value_labels,
  variable_colname = 'psrc_variable',
  value_colname = 'psrc_value',
  value_order_colname = 'psrc_value',
  value_label_colname = "final_label")

dcast(tab, mode_type ~ hhincome_broad, value.var = 'pct') %>% 
  adorn_totals() %>% 
  flextable() %>%
  colformat_double(suffix = '%')

```

# TABLE 31: TRAVEL MODE, BY AGE GROUP (WEIGHTED)

```{r table 31, echo = FALSE, warning = FALSE}

prep = hts_prep_variable(summarize_var = 'mode_type',
                         summarize_by = 'age',
                         variables_dt = variable_list,
                         data = tables,
                         id_cols = ids,
                         wt_cols = wts,
                         missing_values = c("995", "NULL"))


output = hts_summary(prep$cat,
                     summarize_var = 'mode_type',
                     summarize_by = 'age',
                     summarize_vartype = 'categorical',
                     id_cols = ids,
                     wtname = 'trip_weight')

tab = output$summary$wtd

tab$pct = round(tab$prop * 100, 2)

# tab = tab[mode_type != 'NULL']

tab = factorize_df(
  tab,
  value_labels,
  variable_colname = 'psrc_variable',
  value_colname = 'psrc_value',
  value_order_colname = 'psrc_value',
  value_label_colname = "final_label")

dcast(tab, mode_type ~ age, value.var = 'pct') %>% 
  adorn_totals() %>% 
  flextable() %>%
  colformat_double(suffix = '%')

```

# Percentage of trips by trip purpose-------------------------------------------

# TABLE 32: TRAVEL MODE, BY TIME OF DAY (WEIGHTED)

```{r table 32, echo = FALSE, warning = FALSE}

# Create binned var for departure time
tables$trip[, depart_time_binned := fcase(
  depart_time_hour %in% 6:8, 'AM Peak',
  depart_time_hour %in% 9:14, 'Midday',
  depart_time_hour %in% 15:17, 'PM Peak',
  depart_time_hour %in% 18:19, 'Evening',
  depart_time_hour %in% c(0:5, 20:24), 'Night'
)]

# check that assignment is correct
tables$trip[, .N, .(depart_time_hour, depart_time_binned)][order(depart_time_hour)]

# add to variable list
variable_list = rbind(
  variable_list,
  data.table(
    variable = 'depart_time_binned',
    hh = 0,
    person = 0,
    vehicle = 0,
    day = 0,
    trip = 1,
    location = 0,
    description = 'Departure time binned',
    data_type = 'integer/categorical',
    logic = NA,
    is_checkbox = 0,
    shared_name = 'depart_time_binned'
  )
)

prep = hts_prep_variable(summarize_var = 'mode_type',
                         summarize_by = 'depart_time_binned',
                         variables_dt = variable_list,
                         data = tables,
                         id_cols = ids,
                         wt_cols = wts,
                         missing_values = c("995", "NULL"))


output = hts_summary(prep$cat,
                     summarize_var = 'mode_type',
                     summarize_by = 'depart_time_binned',
                     summarize_vartype = 'categorical',
                     id_cols = ids,
                     wtname = 'trip_weight')

tab = output$summary$wtd

tab$pct = round(tab$prop * 100, 2)

# tab = tab[mode_type != 'NULL']

tab = factorize_df(
  tab,
  value_labels,
  variable_colname = 'psrc_variable',
  value_colname = 'psrc_value',
  value_order_colname = 'psrc_value',
  value_label_colname = "final_label")

dcast(tab, mode_type ~ depart_time_binned, value.var = 'pct') %>% 
  adorn_totals() %>% 
  flextable() %>%
  colformat_double(suffix = '%')


```


# TABLE 33: TRIP PURPOSE, BY HOUSEHOLD INCOME (WEIGHTED)

```{r table 33, echo = FALSE, warning = FALSE}

prep = hts_prep_variable(summarize_var = 'dest_purpose_cat',
                         summarize_by = 'hhincome_broad',
                         variables_dt = variable_list,
                         data = tables,
                         id_cols = ids,
                         wt_cols = wts,
                         missing_values = c("995", "NULL"))


output = hts_summary(prep$cat,
                     summarize_var = 'dest_purpose_cat',
                     summarize_by = 'hhincome_broad',
                     summarize_vartype = 'categorical',
                     id_cols = ids,
                     wtname = 'trip_weight')

tab = output$summary$wtd

tab$pct = round(tab$prop * 100, 2)

# tab = tab[dest_purpose_cat != 'NULL']


tab = factorize_df(
  tab,
  value_labels,
  variable_colname = 'psrc_variable',
  value_colname = 'psrc_value',
  value_order_colname = 'psrc_value',
  value_label_colname = "final_label")

dcast(tab, dest_purpose_cat ~ hhincome_broad, value.var = 'pct') %>% 
  adorn_totals() %>% 
  flextable() %>%
  colformat_double(suffix = '%')

```


# TABLE 34: TRIP PURPOSE, BY TIME OF DAY (WEIGHTED)

```{r table 34, echo = FALSE, warning = FALSE}

prep = hts_prep_variable(summarize_var = 'dest_purpose_cat',
                         summarize_by = 'depart_time_binned',
                         variables_dt = variable_list,
                         data = tables,
                         id_cols = ids,
                         wt_cols = wts,
                         missing_values = c("995", "NULL"))


output = hts_summary(prep$cat,
                     summarize_var = 'dest_purpose_cat',
                     summarize_by = 'depart_time_binned',
                     summarize_vartype = 'categorical',
                     id_cols = ids,
                     wtname = 'trip_weight')

tab = output$summary$wtd

tab$pct = round(tab$prop * 100, 2)

# tab = tab[dest_purpose_cat != 'NULL']

tab = factorize_df(
  tab,
  value_labels,
  variable_colname = 'psrc_variable',
  value_colname = 'psrc_value',
  value_order_colname = 'psrc_value',
  value_label_colname = "final_label")

dcast(tab, dest_purpose_cat ~ depart_time_binned, value.var = 'pct') %>% 
  adorn_totals() %>% 
  flextable() %>%
  colformat_double(suffix = '%')

```

# TABLE 35: TRIP PURPOSE, BY DISTANCE (WEIGHTED)

```{r table 35, echo = FALSE, warning = FALSE}

tables$trip[, distance_miles := as.numeric(distance_miles)]

tables$trip[, distance_miles_binned := fcase(
  distance_miles < 1, '< 1 mile',
  distance_miles >= 1 & distance_miles < 2, '1-2 miles',
  distance_miles >= 2 & distance_miles < 4, '2-4 miles',
  distance_miles >= 4 & distance_miles < 6, '4-6 miles',
  distance_miles >= 6 & distance_miles < 8, '6-8 miles',
  distance_miles >= 8 & distance_miles < 10, '8-10 miles',
  distance_miles >= 10 & distance_miles < 12, '10-12 miles',
  distance_miles >= 12 & distance_miles < 14, '12-14 miles',
  distance_miles >= 14 & distance_miles < 16, '14-16 miles',
  distance_miles >= 16 & distance_miles < 18, '16-18 miles',
  distance_miles >= 18 & distance_miles < 20, '18-20 miles',
  distance_miles >= 20, '>= 20 miles'
)]

head(tables$trip[, .(distance_miles, distance_miles_binned)], 20)

# add to variable list
variable_list = rbind(
  variable_list,
  data.table(
    variable = 'distance_miles_binned',
    hh = 0,
    person = 0,
    vehicle = 0,
    day = 0,
    trip = 1,
    location = 0,
    description = 'Distance miles binned',
    data_type = 'categorical',
    logic = NA,
    is_checkbox = 0,
    shared_name = 'distance_miles_binned'
  )
)


prep = hts_prep_variable(summarize_var = 'dest_purpose_cat',
                         summarize_by = 'distance_miles_binned',
                         variables_dt = variable_list,
                         data = tables,
                         id_cols = ids,
                         wt_cols = wts,
                         missing_values = c("995", "NULL"))


output = hts_summary(prep$cat,
                     summarize_var = 'dest_purpose_cat',
                     summarize_by = 'distance_miles_binned',
                     summarize_vartype = 'categorical',
                     id_cols = ids,
                     wtname = 'trip_weight')

tab = output$summary$wtd

tab$pct = round(tab$prop * 100, 2)

# tab = tab[dest_purpose_cat != 'NULL']

tab = factorize_df(
  tab,
  value_labels,
  variable_colname = 'psrc_variable',
  value_colname = 'psrc_value',
  value_order_colname = 'psrc_value',
  value_label_colname = "final_label")

dcast(tab, dest_purpose_cat ~ distance_miles_binned, value.var = 'pct') %>% 
  adorn_totals() %>% 
  flextable() %>%
  colformat_double(suffix = '%')


```

# Travel Day activities---------------------------------------------------------

TABLE 36: SUMMARY OF TELEWORK TIME, BY DAY OF WEEK (EMPLOYED ADULTS, WEIGHTED)

```{r table 36, echo = FALSE, warning = FALSE}

tables$day[, telework_time_binned := fcase(
  telework_time > 0 & telework_time < 60, '0-1 hours',
  telework_time >= 60 & telework_time < 360, '1-6 hours',
  telework_time >= 360, '6+ hours',
  telework_time == 0, 'Did not telework'
)]

# add to variable list
variable_list = rbind(
  variable_list,
  data.table(
    variable = 'telework_time_binned',
    hh = 0,
    person = 0,
    vehicle = 0,
    day = 1,
    trip = 0,
    location = 0,
    description = 'Teleworking time binned',
    data_type = 'categorical',
    logic = NA,
    is_checkbox = 0,
    shared_name = 'telework_time_binned'
  )
)


# get employed values from codebook
employed_values = value_labels[
  psrc_variable == 'employment' & 
    str_detect(
      final_label,
      'Employed full time|Employed part time|Self-employed|volunteer'
    ), 
  psrc_value
]

# filter data to only employed persons
employed_tables = hts_filter_data(tables,
                                  tables$person[employment %in% employed_values, person_id],
                                  'person_id')



prep = hts_prep_variable(summarize_var = 'telework_time_binned',
                         summarize_by = 'travel_dow',
                         variables_dt = variable_list,
                         data = employed_tables,
                         id_cols = ids,
                         wt_cols = wts)


output = hts_summary(prep$cat,
                     summarize_var = 'telework_time_binned',
                     summarize_by = 'travel_dow',
                     summarize_vartype = 'categorical',
                     id_cols = ids,
                     wtname = 'day_weight_2023')

tab = output$summary$wtd

tab$pct = round(tab$prop * 100, 2)

tab = factorize_df(
  tab,
  value_labels,
  variable_colname = 'psrc_variable',
  value_colname = 'psrc_value',
  value_order_colname = 'psrc_value',
  value_label_colname = "final_label")

dcast(tab, telework_time_binned ~ travel_dow, value.var = 'pct') %>% 
  adorn_totals() %>% 
  flextable() %>%
  colformat_double(suffix = '%')


```

# TABLE 37: SUMMARY OF REPORTED DELIVERIES ON TRAVEL DAY (WEIGHTED)

```{r table 37, echo = FALSE, warning = FALSE}

deliver_vars = variable_list[
  shared_name == 'deliver' &
    variable != 'deliver_none',
  variable
]


tables$day[, num_deliver := rowSums(tables$day[, ..deliver_vars] == 1)]

tables$day[, deliver_multiple := as.integer(ifelse(num_deliver > 1, 1, 0))]

tables$day[deliver_multiple == 1, `:=`
           (deliver_elsewhere = 0,
             deliver_food = 0,
             deliver_grocery = 0,
             deliver_office = 0,
             deliver_other = 0,
             deliver_package = 0,
             deliver_work = 0)]

# add to variable list
variable_list = rbind(
  variable_list,
  data.table(
    variable = 'deliver_multiple',
    hh = 0,
    person = 0,
    vehicle = 0,
    day = 1,
    trip = 0,
    location = 0,
    description = 'Deliveries and home services -- Multiple deliveries',
    data_type = 'categorical',
    logic = NA,
    is_checkbox = 0,
    shared_name = 'deliver'
  )
)


prep = hts_prep_variable(summarize_var = 'deliver',
                         variables_dt = variable_list,
                         data = tables,
                         id_cols = ids,
                         wt_cols = wts)


output = hts_summary(prep$cat,
                     summarize_var = 'deliver',
                     summarize_vartype = 'checkbox',
                     checkbox_yesval = 1,
                     id_cols = ids,
                     wtname = 'day_weight_2023')

tab = output$summary$wtd

tab$pct = round(tab$prop * 100, 2)

tab[, .(deliver, pct)] %>% 
  adorn_totals() %>%
  flextable()

```

